---
title: "Benchmark C++ Implementation"
format: 
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    toc-title: "Contents"
    toc-expand: true
    fig-align: center
    number-sections: true
editor: source
author: Hung Pham
execute:
  echo: false
  message: false
  warning: false
---

# Synopsis
Given the refractors, the codes are now modular enough that I've implemented bdlim1 in `{RcppArmadillo}`. This file shows the results of this implementation.

# Method
Fit both bdlim1 in R and in C++, compare the draws of `dlfun`.

```{r}
library(bdlim)
library(posterior)
library(ggplot2)
library(ggdist)

# Helper functions
draws_to_long <- function(draws, i, v) {
  as_draws_df(
    subset_draws(draws, variable = v, regex = TRUE, iteration = i)
  ) |> 
    tidyr::pivot_longer(cols = contains(v))
}

get_dlfun <- function(r_draws, cpp_draws, variable = ".dl", iters) {
  r_fit_dlfun <- draws_to_long(r_draws, iters, variable)
  r_fit_dlfun$fit <- "R"
  
  cpp_fit_dlfun <- draws_to_long(cpp_draws, iters, variable)
  cpp_fit_dlfun$fit <- "C++"
  
  dplyr::bind_rows(r_fit_dlfun, cpp_fit_dlfun)
}

plot_dlfun <- function(d) {
  d |>
    ggplot(aes(
      y = value,
      x = fit,
      color = fit,
      fill = fit
    )) +
    facet_wrap( ~ name, scales = "free_y") +
    stat_slab(alpha = 0.75) +
    stat_pointinterval(position = position_dodge(width = .2, preserve = "single")) +
    scale_color_manual(values = my_cols) +
    scale_fill_manual(values = my_cols) +
    bayesplot::theme_default()
}
```

# Gaussian
```{r}
gaussian_wrap <- function(seed, fit_cpp, nits) {
  set.seed(seed)
  bdlim4(
    y = sbd_bdlim$bwgaz,
    exposure = sbd_bdlim[, paste0("pm25_", 32:37)], # fit only weeks 32:37
    covars = sbd_bdlim[, c("MomPriorBMI"), drop = F],
    group = as.factor(sbd_bdlim$ChildSex),
    model = "n",
    df = 5,
    nits = nits,
    cpp = fit_cpp
  )
}

my_cols <- c("#1B9E77", "#D95F02")
```

For Gaussian, using `{Rcpp}` we get the exact results as the R fit because `{Rcpp}` uses the same seed as R.
```{r}
g_r <- gaussian_wrap(seed = 1234, fit_cpp = FALSE, nits = 200)
g_cpp <- gaussian_wrap(seed = 1234, fit_cpp = TRUE, nits = 200)
```

```{r}
get_dlfun(g_r$fit_n$draws, g_cpp$fit_n$draws, iters = 101:200) |> 
  plot_dlfun()
```

```{r}
benchmark_wrap <- function(nit, wrapper, seed = 123) {
  bm <- lapply(nit, \(x) {
    bench::mark(
      r = wrapper(
        seed = seed,
        fit_cpp = FALSE,
        nits = x
      ),
      cpp = wrapper(
        seed = seed,
        fit_cpp = TRUE,
        nits = x
      ),
      check = F,
      filter_gc = FALSE, 
      relative = TRUE
    ) |> 
      dplyr::mutate(nits = x)
  })
  
  compare_tbl <- do.call(dplyr::bind_rows, bm) |> 
    dplyr::mutate(expression = names(expression)) |> 
    janitor::clean_names()
  
  return(compare_tbl)
}

gaussian_bm <- benchmark_wrap(c(50, 500, 1000), gaussian_wrap)
```

```{r}
bm_plot <- function(bm, what) {
  bm |>
    ggplot(aes(x = nits, y = {{what}}, color = expression)) +
    geom_point() +
    scale_color_manual(values = my_cols) +
    geom_line(aes(group = expression)) +
    theme_bw()
}
```

Next we benchmark. We see that C++ is much more memory efficient (x20). These are relative numbers (folds higher).
```{r}
gaussian_bm |> 
  bm_plot(mem_alloc)
```

Up to 5x faster. Probably faster the higher nits is.
```{r}
gaussian_bm |> 
  bm_plot(median)
```

# Logistics
```{r}
set.seed(1234)
sbd_bdlim$y_binom <- rbinom(nrow(sbd_bdlim), 1, 0.5)
logistic_wrap <- function(seed, fit_cpp, nits) {
  set.seed(seed)
  bdlim4(
    y = sbd_bdlim$y_binom,
    exposure = sbd_bdlim[, paste0("pm25_", 32:37)],
    covars = sbd_bdlim[, c("MomPriorBMI"), drop = F],
    group = as.factor(sbd_bdlim$ChildSex),
    model = "n",
    family = "binomial",
    df = 5,
    nits = nits,
    cpp = fit_cpp
  )
}
```

For logistic, using `{Rcpp}` also get exact same results as R.
```{r}
l_r <- logistic_wrap(seed = 1234, fit_cpp = FALSE, nits = 200)
l_cpp <- logistic_wrap(seed = 1234, fit_cpp = TRUE, nits = 200)

get_dlfun(l_r$fit_n$draws, l_cpp$fit_n$draws, iters = 101:200) |> 
  plot_dlfun()
```

```{r}
logistic_bm <- benchmark_wrap(c(50, 500, 1000), logistic_wrap, seed = 12345)
```

Similar trends to Gaussian, but less of an increase.
```{r}
logistic_bm |> 
  bm_plot(mem_alloc)
```

Up to 2x faster.
```{r}
logistic_bm |> 
  bm_plot(median)
```
